---
title: "Understanding the yield-curve data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding the yield-curve data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`RSYC_models` contains all published RSYC yield curves. Each row describes one
curve for a forest measure, species, area, and source.

```{r}
library(RSYC)
head(RSYC_models[c(
  "model_id", "response", "response_units", "strata_level", "strata_id",
  "scale", "species", "b1", "b2", "b3", "b4"
)])
```

The `response` column describes the forest measure: `agb` is aboveground
biomass in Mg/ha, and `volume` is total volume in m3/ha. Areas can be tiles,
ecozones, ecoprovinces, ecoregions, or ecodistricts. The table includes curves
fitted across Canada and curves fitted within individual regions. The package
functions currently use only the Canada-wide curves.

For ecological areas, `strata_id` gives the full sequence of names from the
ecozone down to the selected level, separated by `/`. `predict_rsyc()` also
accepts the shorter `strata_name` when that name occurs only once for the
chosen forest measure, species, and ecological level:

```{r}
rsyc_strata(
  response = "agb",
  species = "PICE.MAR",
  strata_level = "ecoregion"
)
```

Use `available_rsyc_models()` to see which curves are available without viewing
all of the fields used to fit and check the curves:

```{r}
available_rsyc_models(
  response = "volume",
  species = "PICE.MAR",
  strata_level = "tile"
)
```

`RSYC_models` also contains measures of uncertainty and variation. These values
may support adjustment with local observations in a future version, but
`predict_rsyc()` does not currently use them.
