---
title: "Model parameters and identifiers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model parameters and identifiers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`RSYC_models` is the stable package-facing model table. Each row is one
published curve identified by response, species, spatial stratum, and model
source.

```{r}
library(RSYC)
RSYC_models[c(
  "model_id", "response", "response_units", "strata_level", "strata_id",
  "scale", "species", "b1", "b2", "b3", "b4"
)]
```

The supported responses are `agb` (Mg/ha) and `volume` (m3/ha). Spatial levels
are `tile`, `ecozone`, `ecoprovince`, `ecoregion`, and `ecodistrict`. The model
contract retains both national and regional rows, but the current prediction
and discovery functions expose national models only.

Canonical ecosystem `strata_id` values encode their hierarchy with `/`.
`predict_rsyc()` also accepts the shorter `strata_name` when it is unique for
the requested response, species, and level:

```{r}
rsyc_strata(
  response = "agb",
  species = "PICE.MAR",
  strata_level = "ecoregion"
)
```

Use `available_rsyc_models()` to inspect model metadata without working with
the full coefficient and diagnostic table:

```{r}
available_rsyc_models(
  response = "volume",
  species = "PICE.MAR",
  strata_level = "tile"
)
```

The uncertainty and variance fields in `RSYC_models` are retained for a future
local-calibration interface. They are not currently applied by
`predict_rsyc()`.
