---
title: "Prediction examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(RSYC)
```

## Estimate yield for one species and area

Choose the forest measure, species, stand age, type of area, and area name or
tile ID. The following examples estimate AGB and volume for black spruce in
tile H14 at three stand ages:

```{r}
predict_rsyc("agb", "PICE.MAR", c(20, 60, 120), "tile", "H14")
predict_rsyc("volume", "PICE.MAR", c(20, 60, 120), "tile", "H14")
```

The same function works for individual species and broad groups such as
`coniferous` and `broadleaf`:

```{r}
species_predictions <- tibble::tibble(
  species = c("PICE.MAR", "PINU.BAN", "POPU.TRE", "coniferous", "broadleaf")
) |>
  dplyr::mutate(
    agb_80 = purrr::map_dbl(
      species,
      \(species) predict_rsyc("agb", species, 80, "tile", "H14")
    )
  )

species_predictions
```

## Multiple tiles and species

The following example estimates yield for several combinations of forest
measure, species, stand age, and tile. Each row of `tile_inputs` describes one
combination:

```{r}
tile_inputs <- tidyr::crossing(
  response = c("agb", "volume"),
  species = c("PICE.MAR", "POPU.TRE", "coniferous"),
  age = c(20, 60, 120),
  strata_level = "tile",
  strata_id = c("H14", "F31")
) |>
  dplyr::mutate(
    prediction = purrr::pmap_dbl(
      list(response, species, age, strata_level, strata_id),
      predict_rsyc
    )
  )

tile_inputs
```

Before estimating many combinations, use `rsyc_species()` or
`available_rsyc_models()` to confirm that a published curve is available for
each species and tile.

## Ecological areas

Use the area name when it occurs only once at the selected ecological level:

```{r}
eco_inputs <- tibble::tribble(
  ~response, ~species,   ~age, ~strata_level, ~strata_id,
  "agb",     "PICE.MAR", 80,  "ecozone",
    "Atlantic Maritime",
  "agb",     "PICE.MAR", 80,  "ecoprovince",
    "Appalachian-Acadian Highlands",
  "agb",     "PICE.MAR", 80,  "ecoregion",
    "Appalachians",
  "agb",     "PICE.MAR", 80,  "ecodistrict",
    "Windsor Lowlands"
)

eco_inputs <- eco_inputs |>
  dplyr::mutate(
    prediction = purrr::pmap_dbl(
      list(response, species, age, strata_level, strata_id),
      predict_rsyc
    )
  )

eco_inputs[c("strata_level", "strata_id", "prediction")]
```

Use `rsyc_strata()` to list area names and see where they lie within Canada's
ecological classification. For example, more than one ecodistrict named
`Muskwa` has a curve for this species and forest measure:

```{r}
rsyc_strata("agb", "PICE.MAR", "ecodistrict") |>
  dplyr::filter(strata_name == "Muskwa")

predict_rsyc(
  "agb", "PICE.MAR", 80, "ecodistrict",
  "Boreal Cordillera/Hay-Slave Lowlands/Muskwa Plateau/Muskwa"
)
```

## Several complete curves

This example estimates yield at ages 1 to 150 for each species and area listed
in `curve_inputs`:

```{r}
curve_inputs <- tibble::tribble(
  ~response, ~species,     ~strata_level, ~strata_id,
  "agb",     "PICE.MAR",   "tile",       "H14",
  "agb",     "POPU.TRE",   "tile",       "F31",
  "volume",  "coniferous", "tile",       "N3"
)
curve_inputs <- curve_inputs |>
  dplyr::mutate(
    age = list(1:150),
    prediction = purrr::pmap(
      list(response, species, age, strata_level, strata_id),
      predict_rsyc
    )
  )

curve_inputs

curve_predictions <- curve_inputs |>
  tidyr::unnest(c(age, prediction))

curve_predictions
```

The final table, `curve_predictions`, can then be plotted or used in further
analysis. AGB and volume are shown in separate panels because they use
different units:

```{r curve-plot, fig.width=8, fig.height=5, fig.alt="RSYC yield curves by stand age for three species and tile combinations."}
curve_predictions |>
  dplyr::mutate(
    response = factor(
      response,
      levels = c("agb", "volume"),
      labels = c("AGB (Mg/ha)", "Volume (m3/ha)")
    ),
    curve = paste(species, strata_id, sep = " - ")
  ) |>
  ggplot2::ggplot(ggplot2::aes(age, prediction, colour = curve)) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::facet_wrap(ggplot2::vars(response), scales = "free_y") +
  ggplot2::labs(
    x = "Stand age (years)",
    y = "Predicted yield",
    colour = "Species and tile"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position = "bottom")
```
