---
title: "Prediction examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(RSYC)
```

## Individual predictions

The five arguments are response, species, age, spatial level, and spatial ID.
They can be supplied positionally:

```{r}
predict_rsyc("agb", "PICE.MAR", c(20, 60, 120), "tile", "H14")
predict_rsyc("volume", "PICE.MAR", c(20, 60, 120), "tile", "H14")
```

Species-specific curves and broad model groups use the same interface:

```{r}
species_predictions <- tibble::tibble(
  species = c("PICE.MAR", "PINU.BAN", "POPU.TRE", "coniferous", "broadleaf")
) |>
  dplyr::mutate(
    agb_80 = purrr::map_dbl(
      species,
      \(species) predict_rsyc("agb", species, 80, "tile", "H14")
    )
  )

species_predictions
```

## Multiple tiles and species

Create one row per request and apply the five tibble columns directly to the
five prediction arguments:

```{r}
tile_inputs <- tidyr::crossing(
  response = c("agb", "volume"),
  species = c("PICE.MAR", "POPU.TRE", "coniferous"),
  age = c(20, 60, 120),
  strata_level = "tile",
  strata_id = c("H14", "F31")
) |>
  dplyr::mutate(
    prediction = purrr::pmap_dbl(
      list(response, species, age, strata_level, strata_id),
      predict_rsyc
    )
  )

tile_inputs
```

Before constructing a large request table, use `species_in_tile()` or
`available_rsyc_models()` to verify that every requested combination exists.

## Nested ecological strata

Use the name at the requested ecological level when it identifies one matching
stratum:

```{r}
eco_inputs <- tibble::tribble(
  ~response, ~species,   ~age, ~strata_level, ~strata_id,
  "agb",     "PICE.MAR", 80,  "ecozone",
    "Atlantic Maritime",
  "agb",     "PICE.MAR", 80,  "ecoprovince",
    "Appalachian-Acadian Highlands",
  "agb",     "PICE.MAR", 80,  "ecoregion",
    "Appalachians",
  "agb",     "PICE.MAR", 80,  "ecodistrict",
    "Windsor Lowlands"
)

eco_inputs <- eco_inputs |>
  dplyr::mutate(
    prediction = purrr::pmap_dbl(
      list(response, species, age, strata_level, strata_id),
      predict_rsyc
    )
  )

eco_inputs[c("strata_level", "strata_id", "prediction")]
```

Use `rsyc_strata()` to inspect short names and their canonical paths. For
example, `Muskwa` is ambiguous for this species and response:

```{r}
rsyc_strata("agb", "PICE.MAR", "ecodistrict") |>
  dplyr::filter(strata_name == "Muskwa")

predict_rsyc(
  "agb", "PICE.MAR", 80, "ecodistrict",
  "Boreal Cordillera/Hay-Slave Lowlands/Muskwa Plateau/Muskwa"
)
```

## One complete curve per row

A list-column allows each input row to hold an age vector and its corresponding
yield curve:

```{r}
curve_inputs <- tibble::tribble(
  ~response, ~species,     ~strata_level, ~strata_id,
  "agb",     "PICE.MAR",   "tile",       "H14",
  "agb",     "POPU.TRE",   "tile",       "F31",
  "volume",  "coniferous", "tile",       "N3"
)
curve_inputs <- curve_inputs |>
  dplyr::mutate(
    age = list(1:200),
    prediction = purrr::pmap(
      list(response, species, age, strata_level, strata_id),
      predict_rsyc
    )
  )

curve_inputs

curve_predictions <- curve_inputs |>
  tidyr::unnest(c(age, prediction))

curve_predictions
```

The `age` and `prediction` list-columns can then be expanded or plotted using
the user's preferred data-manipulation and graphics packages.
